---
title: "LendingClubAnalysis"
author: "Steve Isaacs, Kailing See, Sally Guo, Cristian Benavides"
date: "1/12/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Business Problem: Assessing credit worthiness of customers in Lending Club and their probability of defaulting.

Data set: https://www.kaggle.com/wendykan/lending-club-loan-data

### 1.Introduction to the Business Problem

Lending Club is the world’s largest online marketplace connecting borrowers and investors.They operate at a lower cost than traditional bank lending programs and pass the savings on to borrowers in the form of lower rates and to investors in the form of solid returns.
Source: https://www.lendingclub.com/

This project aims to analyse Lending Club's issued loans over a 9 year period (2007-2015) and identify early indicators that could predict a customer's probability of defaulting. The insight obtained could help to drive differentiated credit approval processes for different customer segments (e.g faster approval time, minimal security for low risk customers).

### 2.Process to solve the Business Problem

We would suggest the following process:      
1. Understand Lending Club Business Model & Data Set    
2. Clean the data, identify relevant attributes/columns & separate training and testing dataset    
3. Generate Hypothesis    
4. Apply Dimensionality Reduction on Hypothesis     
5. Apply Segmentation & Classification Tree on Training Dataset    
6. Apply Algorithm on Testing Dataset (Iterate if required)    
7. Capture insights & Summarise    

**OBoulant comments :**
If you want to "identify early indicators that could predict a customer’s probability of defaulting", you have to identify into the dataset, a column that would "represent" the "customer’s probability of defaulting". 
--> Do you have an idea of which column within the data could play this role? ProjectData$loan_status

### 3.Key Summary of Data

```{r eval = FALSE, echo=TRUE, comment=NA, warning=FALSE, message=FALSE,results='markup'}
ProjectData<- read.csv(file = "Data/loan.csv", header = TRUE, sep=",", stringsAsFactors = FALSE, nrows=10000)
# OBoulant : i added the "stringsAsFactors = FALSE", because I want to look at the variables represented by character strings
# SIsaacs: I added "nrows = 10000" to work with a smaller more manageable data set
# print(colnames(ProjectData))
# summary(ProjectData)

ProjectDataClean <- ProjectData$id
ProjectDataClean <- cbind(ProjectDataClean, ProjectData$loan_status, ProjectData$loan_amnt, ProjectData$term, ProjectData$annual_inc, ProjectData$emp_length, ProjectData$pub_rec, ProjectData$open_acc, ProjectData$delinq_2yrs)
colnames(ProjectDataClean) <- c("ID", "LoanStatus", "LoanAmount", "LoanTerm", "AnnualIncome", "EmploymentLength", "PublicRecords", "NumberCreditLines", "DelinquencyPrior2Years")

#Convert the data to numbers
for (r in ProjectDataClean[,"LoanStatus"]){
  if (i=="Charged Off"){
    i <- 1
  }
  else {
  i <- 0
  }
  ProjectDataClean[r,"LoanStatus"]<-i
}

#Scale the data

ProjectDataClean$Loan.Amount <- ProjectDataClean$Loan.Amount - 

#Consider loan grade, subgrade and interest rate separately too

#SIsaacs Questions: how to scale data?                          
                          
# OBoulant comments :
######################

# You should decide what you want to do with categorical variables 
##################################################################

## 1 - term
# Should term be represented as character strings ("36 months" and "60 months") or by 36 and 60 ? 
unique(ProjectData$term)
ProjectData$term <- as.factor(ProjectData$term)
## 2 - Grades 
unique(ProjectData$grade)
ProjectData$grade <- as.factor(ProjectData$grade)
## 3 - sub_grade
unique(ProjectData$sub_grade)
ProjectData$sub_grade <- as.factor(ProjectData$sub_grade)
## 4 - emp_title
unique(ProjectData$emp_title)
head(table(ProjectData$emp_title))
sum(ProjectData$emp_title == '') # Many blanks !
# This field should be re-processed
## 5 - emp_length
unique(ProjectData$emp_length)
ProjectData$emp_length <- as.factor(ProjectData$emp_length)
## 6 - home_ownership
unique(ProjectData$home_ownership)
ProjectData$home_ownership <- as.factor(ProjectData$home_ownership)
## 7 - home_ownership
unique(ProjectData$verification_status)
ProjectData$verification_status <- as.factor(ProjectData$verification_status)
## 8 - issue_d
unique(ProjectData$issue_d)
ProjectData$issue_d <- as.factor(ProjectData$issue_d)
# Maybe should be separated in issue_year and issue_month
## 9 - loan_status
unique(ProjectData$loan_status)
ProjectData$loan_status <- as.factor(ProjectData$loan_status)
## 10 - pymnt_plan
unique(ProjectData$pymnt_plan)
ProjectData$pymnt_plan <- as.factor(ProjectData$pymnt_plan)
## 11 - url
unique(ProjectData$url) # Too many modalities, keep it in character strings - Not to be used !
## 12 - desc
unique(ProjectData$desc) # Too many modalities, keep it in character strings - Not to be used !
## 13 - purpose 
unique(ProjectData$purpose)
ProjectData$purpose <- as.factor(ProjectData$purpose)
## 14 - title
unique(ProjectData$title)
# Either should be re-processed or not to be used ! Or do a word count, i don't know
## 15 - zip_code
unique(ProjectData$zip_code)
ProjectData$zip_code <- as.factor(ProjectData$zip_code)
## 16 - addr_state
unique(ProjectData$addr_state)
ProjectData$addr_state <- as.factor(ProjectData$addr_state)
## 17 - earliest_cr_line
unique(ProjectData$earliest_cr_line)
# Some missing Data
# Maybe should be separated into earliest_cr_line_month and earliest_cr_line_year
ProjectData$earliest_cr_line <- as.factor(ProjectData$earliest_cr_line)
## 18 - initial_list_status
unique(ProjectData$initial_list_status)
ProjectData$initial_list_status <- as.factor(ProjectData$initial_list_status)
## 19 - last_pymnt_d
unique(ProjectData$last_pymnt_d)
ProjectData$last_pymnt_d <- as.factor(ProjectData$last_pymnt_d)
# Some missing Data
# Maybe should be separated into last_pymnt_month and last_pymnt_year
## 20 - next_pymnt_d
unique(ProjectData$next_pymnt_d)
# Some missing Data
# Maybe should be separated into next_pymnt_month and next_pymnt_year
ProjectData$next_pymnt_d <- as.factor(ProjectData$next_pymnt_d)
# 21 - last_credit_pull_d
unique(ProjectData$last_credit_pull_d)
# Some missing Data
# Maybe should be separated into last_credit_pull_month and last_credit_pull_year
ProjectData$last_credit_pull_d <- as.factor(ProjectData$last_credit_pull_d)
# 22 - application_type
unique(Projectdata$application_type)
ProjectData$application_type <- as.factor(ProjectData$application_type)
# 23 - verification_status_joint
unique(ProjectData$verification_status_joint)
ProjectData$verification_status_joint <- as.factor(ProjectData$verification_status_joint)

# Result Dataset
summary(ProjectData)

```

### 4.Dimensionality Reduction


### 5.Clustering and Segmentation

### 6. Results 

>>>>>>> 7664b6275553d5ba40165c41ff18eb644596ceb4
