---
title: "LendingClubAnalysis"
author: "Steve Isaacs"
date: "1/12/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Business Problem: Assessing credit worthiness of customers in Lending Club and their probability of defaulting.

Data set: https://www.kaggle.com/wendykan/lending-club-loan-data

### 1.Introduction to the Business Problem

Lending Club is the world’s largest online marketplace connecting borrowers and investors.They operate at a lower cost than traditional bank lending programs and pass the savings on to borrowers in the form of lower rates and to investors in the form of solid returns.
Source: https://www.lendingclub.com/ 

THis project aims to analyse Lending Club's issued loans over 9 year period (2007-2015) and zdentify early indicators that could predict a customer's probability of defaulting. The insight obtained could help to drive differentiated credit approval process for different customer segment (e.g faster approval time, minimal security for low risk customers)

<br>
### 2.Process to solve the Business Problem

We would suggest the following process
1. Understand Lending Club Business Model & Data Set
2. Clean the data, identify relevant attributes/columns & separate training and testing dataset
3. Generate Hypothesis 
4. Apply Dimensionality Reduction on Hypothesis
5. Apply Segmentation & Classification Tree on Training Dataset
6. Apply Algorithm on Testing Dataset (Iterate if required)
7. Capture insights & Summarise

<br>

### 3.Key Summary of Data

```{r eval = FALSE, echo=TRUE, comment=NA, warning=FALSE, message=FALSE,results='markup'}
Projectdata<- read.csv(file = "Data/loan.csv", header = TRUE, sep=","))
print(colnames(Projectdata))
summmary(Projectdata)
```
<br>

### 4.Dimensionality Reduction


<br>
### 5.Clustering and Segmentation

<br>
### 6. Results 

